``Colloidal cold rain'' as a novel route to crystallization via viscoelastic phase separation

### 2015-10-13 crystal-gel/science/SOM2.tex
Corrected the SOM2

\noindent
{\bf Experimental.}
We use \textsc{pmma} (poly(methyl methacrylate)) colloids sterically stabilized with methacryloxypropyl terminated \textsc{pdms}(poly(dimethyl siloxane)) and fluorescently labelled with rhodamine isothiocyanate chemically bonded to the \textsc{pmma}. 
The average colloid diameter is 2.1 $\mu$m. 
We assess that the size distribution of our particle is Gaussian with a polydispersity below 5\% via direct confocal measurements~\cite{Leocmach2013}.
This small polydispersity allows crystallization.
We disperse the particles in refractive index and density matching mixture of cis-decalin (Tokyo Kasei) and bromocyclohexane (Sigma-Aldrich).

To induce short-ranged depletion attraction, we use polystyrene (TOSOH) of molecular weight 3.8 Mdalton as non-adsorbing polymer. 
Experiments are conducted at 27 $^\circ$C, some 80 $^\circ$C above the theta temperature in this solvents mixture~\cite{Royall2007}. A Flory scaling of the measurements of~\cite{lu2008gelation} yields a radius of gyration $R_g=76$ nm, and thus the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.07$.

In the absence of salt, the Debye length is expected to reach several $\mu$m and the (weakly) charged colloids experience a long range electrostatic repulsion. We confirm that colloids never come close enough to feel the short-ranged attraction. Screening by tetrabutylammonium bromide (Fluka) at saturated concentration brings down the Debye length to about 100 nm, practically discarding the repulsion. 
Thus, salt injection can screen the Coulomb repulsion and make the polymer-induced depletion attraction effective, initiating phase separation.

@@

The data are collected on a Leica SP5 confocal microscope, using 532 nm laser excitation. The scanning volume is 98 $\times$ 98 $\times$ 53 $\mu$m$^3$, which contains $\sim 10^4$ colloid particles.  The Brownian time is $\tau_B \approx 4.5$ s. To be able to follow individual trajectories, we perform a 3D scan every 10 s at early time and every 30 s later.

@@

\noindent
{\bf Structural analysis.}
To account for imprecision in particle localisation close to contact, we consider two particles bonded if their center-to-center distance is within 2.3 $\mu$m $>\sigma+2R_g$.

@@

>>>>

\noindent
{\bf Experiments}

We use \textsc{pmma} (poly(methyl methacrylate)) colloids sterically stabilized with methacryloxypropyl terminated \textsc{pdms}(poly(dimethyl siloxane)) and fluorescently labelled with rhodamine isothiocyanate chemically bonded to the \textsc{pmma}. 
The average colloid diameter is 2.2 $\mu$m, measured via the diffusion constant in dilute conditions without polymer. The Brownian time is $\tau_B \approx 2.3$ s. 
We assess that the size distribution of our particle is Gaussian with a polydispersity below 5\% via direct confocal measurements~\cite{Leocmach2013}.
This small polydispersity allows crystallization.
We disperse the particles in refractive index and density matching mixture of cis-decalin (Tokyo Kasei) and bromocyclohexane (Sigma-Aldrich). 
The precision of density matching is $\sim 10^{-4}$ mg/ml, for which the gravity effect on our gelation processes is negligible up to 12 hours.

To induce short-ranged depletion attraction, we use polystyrene (TOSOH) of molecular weight 3.8 Mdalton as non-adsorbing polymer. 
Experiments are conducted at 27 $^\circ$C, some 80 $^\circ$C above the theta temperature in this solvents mixture~\cite{Royall2007}. A Flory scaling of the measurements of~\cite{lu2008gelation} yields a radius of gyration $R_g=82$ nm, and thus the polymer-colloid size ratio is $q_R=2R_g/\sigma=0.075$ and the overlap mass fraction of polymer $2.1$~mg/g.

In the absence of salt, the Debye length is expected to reach a few $\mu$m and the (weakly) charged colloids experience a long range electrostatic repulsion. We confirm that colloids never come close enough to feel the short-ranged attraction. Screening by tetrabutylammonium bromide (Fluka) at saturated concentration brings down the Debye length to about several nanometers practically discarding the repulsion. 
Thus, salt injection can screen the Coulomb repulsion and make the polymer-induced depletion attraction effective, initiating phase separation.

@@

The data are collected on a Leica SP5 confocal microscope, using 532 nm laser excitation. The scanning volume is 98 $\times$ 98 $\times$ 53 $\mu$m$^3$, which contains $\sim 10^4$ colloid particles. Our spatial resolution is 192 nm / pixel. To be able to follow individual trajectories, we perform a 3D scan every 10~s ($\approx 4\tau_B$) at early time and every 30 s later.

@@

\noindent
{\bf Structural analysis}

To account for imprecision in particle localisation close to contact, we consider two particles bonded if their center-to-center distance is within 2.3 $\mu$m $>\sigma+2R_g$.

@@

\noindent
{\bf Supplementary Movies}

\emph{Supplementary Movie 1}. This movie shows the early stage of gelation process of $\phi\approx 0.30$ and $c_p=0.38$ mg/g, cropped from the original confocal images at the same z-position as Fig.~3A with scanning duration of 1160 s (~580~$\tau$). We quickly started scanning the sample after exchanging salt-reservoir with the solvent including salt. At the beginning, salt had not reached the scanning volume and colloids were dispersed by electrostatic repulsion. At 0:02 in the movie, salt diffused into the scanning volume and screened the surface charges of colloids. Then, colloids immediately initiated phase separation by short-ranged attractions.

\emph{Supplementary Movie 2}.
This movie shows the late stage after Supplemental Movie 1 with scanning duration of 27000~s (~13500$\tau$).
